---
title: Updated Mysis 2022 Lake Michigan Data Exploratory Plots
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(rmarkdown)
library(mgcv)
library(sf)
library(raster)
```
```{r}
GLNPO <- read.csv("C:\\Users\\sarah\\Documents\\Cornell\\Mysid Data\\Exploratory Plots\\Mysis-Plots\\MI_OutputByStation_CSV.csv")
summary(GLNPO)
```
```{r}
colnames(GLNPO)
```

Does density increase with depth in Lake Michigan?
```{r}
ggplot(data = GLNPO, mapping = aes(x = Station_Depth_m, y = AvgDens)) + 
  geom_point(mapping = aes(color = Year_), shape = 1) + 
  geom_smooth(method = "lm", se = F)
```

How does density relate to biomass?
```{r}
ggplot(data = GLNPO, mapping = aes(x = AvgBmss_mg, y = AvgDens)) + geom_point(mapping = aes(color = Year_), shape = 1) + geom_smooth(method = "lm", se = F)
```
Does the relationship between biomass and density change year to year?
```{r}
ggplot(data = GLNPO, mapping = aes(x = AvgBmss_mg, y = AvgDens)) + geom_point(mapping = aes(color = Season), shape = 1) + geom_smooth(method = "lm", se = F) + facet_wrap(~Year_)
```

```{r}
ggplot(data = GLNPO, mapping = aes(x = AvgBmss_mg, y = AvgDens)) + geom_point(mapping = aes(color = Season), shape = 1) + geom_smooth(method = "lm", se = F) + facet_wrap(~Season)
```

Density and biomass over the years, colored by season

```{r}
 ggplot(data = GLNPO) + geom_point(mapping = aes(x = Year_, y = AvgDens, color = Season, size = AvgBmss_mg), shape = 9) 
```
```{r}
ggplot(data = GLNPO) + geom_point(mapping = aes(x = Year_, y = AvgDens, color = Season, size = AvgBmss_mg), shape=1) + geom_smooth(mapping = aes(x = Year_, y = AvgDens)) + facet_wrap(~Season)
```
```{r}
ggplot(data = GLNPO) + geom_point(mapping = aes(x = Year_, y = AvgDens, color = Season)) + geom_smooth(mapping = aes(x = Year_, y = AvgDens)) + facet_wrap(~Season)
```
```{r}
ggplot(data = GLNPO) + geom_point(mapping = aes(x = Year_, y = AvgBmss_mg, color = Season)) + geom_smooth(mapping = aes(x = Year_, y = AvgBmss_mg)) + facet_wrap(~Season)
```
```{r}
ggplot(GLNPO, aes(x = Year_, y = AvgDens, color = Season, shape = Season)) + geom_point() + geom_smooth(aes(linetype = Season)) + facet_wrap(~Season)
```
Basic Density and Biomass GAMs

```{r}
Dens <- GLNPO$AvgDens
Year <- GLNPO$Year_
DensGAM <- gam(Dens ~ s(Year))
summary(DensGAM)
```
```{r}
plot(DensGAM, bg="green", col="#336666", xlab = "Year", ylab = "Average Density", main = "Mysis Density in Lake Michigan 2006-2021 Using GLNPO Data")
```

```{r}
Bmss <- GLNPO$AvgBmss_mg
BmssGAM <- gam(Bmss ~ s(Year))
summary(BmssGAM)
```
```{r}
plot(BmssGAM, col="blue", xlab = "Year", ylab = "Average Biomass", main = "Mysis Biomass in Lake Michigan 2006-2021 Using GLNPO Data")
```
Exploratory non-parametric regressions using ggplot

```{r}
GLNPO <- read.csv("C:\\Users\\sarah\\Documents\\Cornell\\Mysid Data\\Exploratory Plots\\Mysis-Plots\\MI_OutputByStation_CSV.csv")
```


```{r}
require(smooth)
library(tidyverse)
library(ggpubr)
```
```{r}
Depth <- GLNPO$Station_Depth_m
Season <- GLNPO$Season
```

Linear regression

```{r}
gg.lm <- ggplot(GLNPO,aes(x = Year_,y = AvgDens, col=Season)) + geom_point() + geom_smooth(method = 'lm') + facet_wrap(~Season)
print(gg.lm)

gg.lm <- ggplot(GLNPO,aes(x = Year_,y = AvgDens)) + geom_point(mapping = aes(size = Depth), shape = 1, color = "darkcyan") + geom_smooth(method = 'lm') + facet_wrap(~Season)
print(gg.lm)
```

Quadratic regression

```{r}
gg.qua <- ggplot(GLNPO,aes(x = Year_, y=AvgDens, col=Season)) + geom_point() + geom_smooth(method = 'lm', formula = y~poly(x,2)) + facet_grid(~Season)
print(gg.qua)
```

Cubic regression

```{r}
gg.cub <- ggplot(GLNPO,aes(x = Year_, y = AvgDens, col=Season)) + geom_point() + geom_smooth(method='lm', formula = y ~ poly(x,2)) + facet_wrap(~Season)
print(gg.cub)
```

Loess

```{r}
gg.lo <- ggplot(GLNPO, aes(x = Year_, y=AvgDens, col=Season)) + geom_point() + geom_smooth(method = 'loess') + facet_wrap(~Season)
print(gg.lo)
```

Natural splines, GAMs

```{r}
gg.ns2 <- ggplot(GLNPO, aes(x = Year_, y = AvgDens, col=Season)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ splines::ns(x,2)) +
  facet_wrap(~Season)
print(gg.ns2)
```
```{r}
gg.ns3 <- ggplot(GLNPO, aes(x = Year_, y = AvgDens, col = Season)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ splines::ns(x,3)) +
  facet_wrap(~Season)
print(gg.ns3)
```
```{r}
gg.bs2 <- ggplot(GLNPO, aes(x = Year_, y = AvgDens, col=Season)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ splines::bs(x,2)) +
  facet_wrap(~Season)
print(gg.bs2)
```
```{r}
gg.ns30 <- ggplot(GLNPO, aes(x = Year_, y = AvgDens, col=Season)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ splines::ns(x,30)) +
  facet_wrap(~Season)
plot(gg.ns30)
```
```{r}
gg.bs30 <- ggplot(GLNPO, aes(x = Year_, y = AvgDens, col=Season)) +
  geom_point() +
  geom_smooth(method = 'gam', formula = y ~ splines::bs(x,30)) +
  facet_wrap(~Season)
plot(gg.bs30)
```
Significance of Station Depth

```{r}
ggplot(data = GLNPO, mapping = aes(x = Station_Depth_m, y = AvgDens)) + geom_point() + geom_smooth(method = "lm", se = F)

summary(lm(AvgDens~Station_Depth_m, data=GLNPO))
```

```{r}
ggplot(data = GLNPO, mapping = aes(x = AvgOfLatitude, y = AvgDens)) + geom_point() + geom_smooth(method = "lm", se = F)
```

```{r}
mi <- st_read("C:\\Users\\sarah\\Documents\\Cornell\\Mysid Data\\Bathymetry_Data\\Lake_Michigan_Contours\\Lake_Michigan_Contours.shp")

mi_coords <- do.call(rbind, st_geometry(mi)) %>% as_tibble() %>% setNames(c("lon", "lat"))
```

